% Lead: Leanne, reviewed by James Mullaney
\section{Summary and Future Releases
\label{sec:summary}}

Rubin Data Preview 1 (DP1) offers an initial look at the first on-sky data products and access services from the Vera C. Rubin Observatory. DP1 forms part of Rubin's Early Science Program, and provides the scientific community with an early opportunity to familiarize themselves with the data formats and access infrastructure for the forthcoming Legacy Survey of Space and Time (LSST).
This early release has a proprietary period of two years, during which time it is  available to Rubin data rights holders only via the cloud-based Rubin Science Platform (RSP). 

In this paper we have described the completion status of the observatory at the time of data acquisition, the commissioning campaign that forms the basis of DP1, and the processing pipelines used to produce early versions of data products.
We provide details on the data products, their characteristics and 
known issues, and describe the RSP. 

The data products described in this paper derive from observations obtained by LSSTComCam. LSSTComCam contains only around 5\% the number of CCDs as the full LSST Science Camera (LSSTCam), yet the DP1 dataset that it has produced will already enable a very broad range of science. 
At \sizeinbytes in size, DP1 covers a total area of \totalarea and contains \nexposures single-epoch images, \ndeepcoadds deep coadded images, \nobjects distinct astrophysical objects, including  \nnewasteroiddiscoveries  new asteroid discoveries. 

While some data products expected from the LSST are not yet available, e.g. cell-based coadds, several others have been provided in DP1 that will not be available in future releases.
Difference images are included in DP1, but in future releases, these will be generated on-demand via services, rather than being provided as pre-produced products. 
The inclusion of these images in DP1 is possible due to the small dataset size, which makes it feasible to include them at this stage. 
As future releases will involve much larger datasets, this approach will no longer be possible.

The RSP is continually under development, and new functionality will continue to be deployed incrementally as it becomes available, and independent of future data releases. 
For example, user query history capabilities, context-aware documentation and a bulk cutout services are just a few of the services currently under development. 

Coincident with the release of DP1, Rubin Observatory begins its Science Validation Surveys with the LSST Science Camera. 
This final commissioning phase will produce a dataset that will form the foundation for the second Rubin Data Preview, DP2, expected around mid to late 2026.
Full operations -- marking the start of the LSST -- is expected to commence by the end of 2025.